---
title: "Quantitative Paper"
author: "Lauryl Zenobi"
date: "`r Sys.Date()`"
output: html_document
bibliography: quantbib.bib
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, cache=TRUE, echo=FALSE, tidy=FALSE)
```

#Introduction
For the purpose of this quantitative project I will explore the theoretical frameworks associated with several key words (see below). The data derived from this exploration will be used to improve upon the background and theoretical framework of my dissertation proposal, which namely theories about analyzing ceramics, interpreting standardization of ceramic industries, and the processes that cause goods (especially staple foods) to become commodities. There is also the potential that this data will provide methodological approaches to the quantitative portion of my dissertation research, which seeks to explore the relationship between the standardization of ceramic ovens used in processing sago starch in Maluku, Indonesia, and the resulting commodification of sago in the context of the spice trade in the Banda Islands. It is useful to understand how standardization, commodification, and ceramics are discussed in the literature, in order to enrich modeling and hypothesis generation. 

#Key Terms and Hypotheses
The questions that are guiding this paper consist of the following: What is the relationship between **standardization** and **ceramics** in the archaeological literature? What is the relationship between **ceramics** and **commodities**? Has the use of these terms changed over time? What theoretical frameworks and models support these types of analyses? 

Due to the small sample size that is expected for these topics and study area, this project will consists of two parts. The first is a statistical analysis of the frequency of the terms identified in the preceding questions. The second part will consist of a close reading of selected articles of interest. The low numbers expected from the statistical analysis necessitate the close readings, which will be beneficial in answering higher level questions about the theoretical frameworks that were utilized in the articles of interest. 

I hypothesize that standardization and ceramics will have the highest correlation, as there is a long history of looking at these topics in archaeological literature. I also suspect that commmodification and standardization will be significantly correlated. However, I do not expect to find a high degree of correlation between ceramics and commodification, as ceramics are generally utilitarian goods that were not likely to have been directly involved in luxury trade networks. 

```{r Data_ingest}
library(JSTORr)
#multiple_archives <- JSTORr ::: JSTOR_unpack_multiple_archives(mydir = "Data")
#save(multiple_archives, file="Data/multiple_archives.RData")
load("Data/multiple_archives.RData")
```

#Exploratory Data Analysis
Statistical analyses were conducted to determine the prevalence of each keyword by decade. Preliminary testing of the questions mentioned above was necessary to determine which research avenue was the most likely to produce information useful for the purpose of this study. The first task was to establish the frequency of key words that are used in the guiding research questions. **Commodity**, **Standardization**, and **Ceramics** were analyzed for their correlation over time. 

```{r}
Figure_1 <- JSTOR_2wordcor(multiple_archives, "commodity", "standardization")
Figure_1$plot
```  
  
Figure 1. Correlation over time between "commodity" and "standardization".

```{r}
Figure_2 <- JSTOR_2wordcor(multiple_archives, "commodity", "ceramics")
Figure_2$plot #2006 is a high year
```  
  
Figure 2. Correlation over time between "commodity" and "ceramics".

```{r}
Figure_3 <- JSTOR_2wordcor(multiple_archives, "standardization", "ceramics")
Figure_3$plot
```  
  
Figure 3. Correlation over time between "standardization" and "ceramics".
  
The results of these plots indicate that while there may not be a large number of instances in which these sets of words appear together in the literature, each plot demonstrates that of the instances in which they do appear, the sets of words are strongly correlated. 

In order to identify publications of interest, each keyword, and variants of each keyword were then analyzed for their frequency over time. No synonym was utilized for "commodity" as that word is very specific and unlikely to produce erroneous results, but "standard" and "standardization" were compared to determine which keyword was most relevant to this study. As expected, "standard" occurred with high proportion. The article in which it appears with the highest proportion was related to mathematical modeling (i.e. mathematical standard errors) and was determined to not be useful for this paper. Standardization, however, produced a highly useful article, which will be discussed in the selected readings section below. 

```{r 1word_commodity}
Figure1word_commodity<- JSTOR_1word(multiple_archives, "commodity")
Figure1word_commodity$plot

commodity_tbl <- Figureoneword$word_by_year
commodity_tbl[which.max(commodity_tbl$word_ratio),]
#url: http://www.jstor.org/stable/279900  use last digits of doi to get to link
```  
  
Figure 4. Frequency of "Commodity" over time.   

```{r 1word_standardization}
Figure1word_standardization<- JSTOR_1word(multiple_archives, "standardization")
Figure1word_standardization$plot
standardization_tbl <- Figure1word_standardization$word_by_year
standardization_tbl[which.max(standardization_tbl$word_ratio),]
```  
  
Figure 5. Frequency of "Standardization" over time.

"Ceramics" and "ceramic" were also compared to determine which keyword would be most productive. Interestingly, the two words produced markedly different articles, with ceramics producing articles that covered a wide variety of topics related to ceramic analyses, including temper analyses and studies on the physical properties of the clay materials. The wide range of topics was not suitable for this study, which seeks to pinpoint theoretical frameworks. Therefore, "ceramic" was utilized as the keyword of interest. "Pottery" was also explored, and the results were more tailored than those of "ceramics", however, the article with the highest proportion of the term was a response article, and not useful for exploring theoretical concepts or methods of explanation. 

```{r 1word_ceramic}
Figure1word_ceramic<- JSTOR_1word(multiple_archives, "ceramic")
Figure1word_ceramic$plot
ceramic_tbl <- Figure1word_ceramic$word_by_year
ceramic_tbl[which.max(ceramic_tbl$word_ratio),] #Sharer and Gifford
```  
  
Figure 6. Frequency of "ceramic" over time. 

```{r standard_frequency_counts, results='hide'}

# top words per decade in articles containing 'X'
standardization_nouns <- JSTOR_dtmofnouns(multiple_archives, word = 'standardization', sparse = 1, POStag = TRUE)

#remove noisy words
custom_stopwords <- c('archaeology', 'university', 'research', 'evidence', 'journal', 'world', 'site', 'cambridge', 'archaeol', 'area', 'region', 'period', 'analysis', 'anthropology', 'springer', 'production', 'figure', 'work', 'world', 'concept', 'human', 'middle', 'altamira', 'culture', 'record', 'citation', 'discipline', 'author', 'proportion', 'literature', 'report', 'approach', 'cambridge')
standardization_freqwords <- JSTOR_freqwords(multiple_archives, standardization_nouns, custom_stopwords, n = 3)
```  
  
Figure 7. Words associated with "standardization" over time. 

```{r standardization_frequency_counts, results='hide'}
standard_nouns <- JSTOR_dtmofnouns(multiple_archives, word = 'standard', sparse = 1, POStag = TRUE)
standard_freqwords <- JSTOR_freqwords(multiple_archives, standard_nouns, custom_stopwords, n = 3)
```  

Figure 8. Words associated with "standard" over time. 


It's clear that "standard" and "standardization" produce two different frequecy plots. From 2008-2010, pottery is the fourth word most commonly printed in association with "standard". It is also a common word during the periods of 1996-1998, the most common word associated with "standard" in the 1993-1995 literature, and the fifth most common word in articles published between 2006-2007. Standardization presents an even clearer image, in 1999-2001 article, pottery is the most common word associated with "standardization", though it falls out of favor in later years. This demonstrates that pottery and standardization/standard regularly are associated in the literature data set.

```{r commodity_frequency_counts, results='hide'}
# top words per decade in articles containing 'X'
commodity_nouns <- JSTOR_dtmofnouns(multiple_archives, word = 'commodity', sparse = 1, POStag = TRUE)

commodity_freqwords <- JSTOR_freqwords(multiple_archives, commodity_nouns, custom_stopwords, n = 3)
```    
  
Figure 9. Words associated with "commodity" over time. 

Pottery and commodity only appear in association within articles published between 1994-1996, however, during this time period, pottery is the second most common word associated with commodity, indicating the two appear at least 400 times in conjunction. Also useful for this project is the literature published between 2006-2008, in which trade and exchange are the two most commonly associated with commodity. The question that is driving this research assumes that commodities are transported as a result of increased trade and exchange, contrary to the expectations of my hypotheses.

```{r pottery_frequency_counts, results='hide'}
# top words per decade in articles containing 'X'
pottery_nouns <- JSTOR_dtmofnouns(multiple_archives, word = 'pottery', sparse = 1, POStag = TRUE)

pottery_freqwords <- JSTOR_freqwords(multiple_archives, pottery_nouns, custom_stopwords, n = 3)
```
  
Figure 10. Words associated with "pottery" over time.

```{r ceramics_frequency_counts, results='hide'}
ceramics_nouns <- JSTOR_dtmofnouns(multiple_archives, word = 'ceramics', sparse = 1, POStag = TRUE)

ceramics_freqwords <- JSTOR_freqwords(multiple_archives, ceramics_nouns, custom_stopwords, n = 3)
```  
  
Figures 11. Words associated with "ceramics" over time. 

The results of using "pottery" show that, even taking out noisey words, the data is very jumbled. The results of a frequency analysis for "ceramics" are relatively similar, possible even more jumbled. It is interesting to note that in both cases, ceramics and pottery produce too many associations to be useful, therefore, for the purpose of the paper, standardization/standard, and commodity will be utilized because they provide much more constrained data sets and are theoretically specific. 

#Selected Readings and Theoretical Frameworks
The exploratory research produced several articles of interest. Each article demonstrates the types of theoretical frameworks that can be applied to understanding commodities and standardization, particularly in reference to the standardization of ceramics. 

Zeitlan 1982 is characteristic of the growth in the use of the keyword **commodity** in archaeological studies. It is clear from the preliminary data exploration that this topic has increased significantly since the 1970s (Figure 4), likely as a result of the growth of processual archaeology [@Zeitlin_1982]. This article discusses commodities in terms of distance-decay models and commodity distribution models. Distance-decay models hypothesize that the abundance of a commodity will decrease proportionate to its distance from source. Commodity distribution models similarly predict the relative scarcity or abundance of non-local goods throughout the trade network [@Zeitlin_1982]. Both types of models suffer from being overly simplistic or mechanistic, without clarifying the actual mechanisms of trade and ignoring the more interesting and complex questions of the relationships between long distance trade and sociopolitical/administrative systems, which are criticisms typical of mechanistic or Causal Nexus models [@Glennan_2010], [@Salmon_1997]. 

Sharer (1970) is an article that typifies the discussion of **ceramics** during the 1970s (see Figure 6), as it focuses on processual approaches towards ceramic analyses - namely a type-variety approach in which ceramic assemblages are categorized into various 'types' in order to answer questions about interaction and trade. This study looked at ceramics from three sites in western El Salvador as a proxy for cultural contact, specialization, trade, and interaction [@Sharer_1970]. 

A discussion of **standardization** is presented in [@Blackman_1993], which is an excellent example of the modestly increasing focus on standardization in recent years within the archaeological literature (see Figure 5). Blackman et al. 1993 define standardization as the increased homogeneity of an artifact type, best observed by comparing two or more assemblages from the same cultural tradition. The reduction in variability within a ceramic assemblage is assumed to be a result of centralized or specialized mass production, in which a desire for cost and labor efficiency result in a selection towards a standardized ceramic form. The standardization hypothesis predicts that standardization of a ceramic assemblage will be present a narrow range of clay materials utilized and a reduction in manufacturing method variability [@Blackman_1993]. This expectation is borne out in the ceramic waster assemblage recovered from Tell Leilan, in Syria, which was studied by [@Blackman_1993]. The results of that study indicate that standardized ceramic production over a long period of time (two to three centuries) will result in some variability to the ceramic assemblage, but the overall expectations of the standardization hypothesis remain true. 

#Discussion of Results
As expected, standardization and ceramics are strongly correlated, with **pottery** being one of the top three words associated with **standardization** in the literature (see Figure 7). However, my second hypothesis that **ceramics** (or its variant, **pottery**) would not be correlated with **commodity** was disproven by the results of the frequency association (Figure 9). This is counter to my expectation that pottery or ceramics would not be associated with luxury trade networks. It is possible, and the targeted reading corroborates this, that ceramics or ceramic production methods are incidentally transferred as a result of any trade network. Lasly, my hypothesis that **standardization** and **commodification** would be strongly correlated was proven correct, as there are many instances in which the two are significantly associated (see Figure 1), and there seems to be a simultaneous trend towards an increasing frequency in both keywords (see Figures 4 and 5).

The results of the quantitative exploratory data and the targeted readings demonstrates the utility of the text mining technique's ability to target articles that are immediately relevant to studies of standardization and commodity exchange. The quantitative data indicates a general trend towards exploring theories of commodity, specialization, and standardization overall, and a trend towards incorporating these theories into interpreting ceramics assemblages. This trend is likely a result of an overall increase in rescue archaeology and an interest in typological studies on the tail of processual archaeology. The application of these theories to ceramic studies indicates that the theory and methodology of this line of inquiry is growing and drawing from a diverse range of theories, including economics, Marxism, scales of economy, and socioeconomic systems. 

#References 

