---
title: "Quantitative Paper"
author: "Lauryl Zenobi "
date: "May 15, 2015"
output: rmarkdown::tufte_handout
---

For the purpose of this quantitative project I want to explore the relationship between several sets of words. This quantitative data will be used to improve upon the background and theoretical framework of my dissertation proposal. There is also the potential that this data will provide methodological approaches to the quantitative portion of my dissertation research. 

Questions that I think would be well-suited to this type of research are listed below.

1. What is the relationships between standardization and ceramics in the archaeological literature? Has this relationship changed over time? What theoretical frameworks support these types of analyses?
2. What is the relationship between sago palm and ceramic analyses in Indonesia? Specifically, what is the statistical relationship between the use of terms such as sago, sago starch, ceramics, and forna (ceramic ovens) in the archaeological literature?
3. What is the relationship between theories about colonial contact and acculturation and changes in ceramic assemblages? Specifically for Indonesia, how do these topics present in the archaeological literature?
4. How has the spice trade and the sago trade been analyzed in the archaeological literature? What is the relationship between terms such as spice trade, nutmeg, cloves, and sago? 


Due to the small sample size that is expected for these topics and study area, this project will consists of two parts. The first is a statistical analysis of the frequency of the terms identified in the preceding questions. The second part will consist of a close reading of selected articles of interest. The low numbers expected from the statistical analysis necessitate the close readings, which will be beneficial in answering higher level questions about the theoretical frameworks that were utilized in the articles of interest. 


```{r Data_ingest, eval=FALSE}
#library(JSTORr)
#multiple_archives <- JSTORr ::: JSTOR_unpack_multiple_archives(mydir = "Data")
#save(multiple_archives, file="Data/multiple_archives.RData")
load("Data/multiple_archives.RData")
```

Codes from the JSTORr package will be utilized for the statistical analysis. 
Preliminary testing of the questions above is necessary to determine which research avenue is the most likely to have a sufficiently sized data set and to produce the most useful information for the purpose of this study. I will start this process with the first potential research question (see above). The first task is to establish the frequency of key words that are used in the first research question. "Commodity", "Standardization", and "Ceramics" were analyzed for their correlation over time. 

```{r, fig.width=4, fig.height=5, message=F, eval=FALSE}
Figure_1 <- JSTOR_2wordcor(multiple_archives, "commodity", "standardization")


Figureoneword <- JSTOR_1word(multiple_archives, "commodity")
commodity_tbl <- Figureoneword$word_by_year
commodity_tbl[which.max(commodity_tbl$word_ratio),]
#url: http://www.jstor.org/stable/279900  use last digits of doi to get to link
```
Figure 1 shows the correlation over time between the words "commodity" and "standardization".

```{r, fig.width=4, fig.height=5, message=F}
Figure_2 <- JSTOR_2wordcor(multiple_archives, "commodity", "ceramics")
```
Figure 2 shows the correlation over time between the words "commodity" and "ceramics".

```{r, fig.width=4, fig.height=5, message=F}
Figure_3 <- JSTOR_2wordcor(multiple_archives, "standardization", "ceramics")
```
Figure 2 shows the correlation over time between the words "standardization" and "ceramics".
  
The results of these plots indicate that while there may not be a large number of instances in which these sets of words appear together in the literature, each plot demonstrates that of the instances in which they do appear, there are a number the sets of words are strongly correlated. This suggests, and close readings can verify this, that the data set for this research question would be rich enough to justify a quantitative method supplemented by a close readings of selected articles. Therefore, question #1 - What is the relationships between standardization and ceramics in the archaeological literature? Has this relationship changed over time? What theoretical frameworks support these types of analyses? - will be the focus of this quantative research paper. 

To determine how each key word of interest (commodity, standardization, and ceramics/pottery) cluster within the literature, a principal component analysis of each terms was conducted. The literature data was first compiled into a vector containing only nouns, as to eliminate erroneous words. Each word was then analyzed against this vector to determine what contexts the word is utilized in within the literature. Outlier and discrete clusters of of the most interest to this research, as those will be the articles in which the terms are used uniquely. 

```{r cluster analysis}
nouns <-JSTOR_dtmofnouns(multiple_archives)
#save(nouns, file="Data/nouns.RData")

Figure_4 <- JSTOR_clusterbywords(nouns, "commodity")

#correlation of two vectors within range of decades
JSTOR_2wordcor(multiple_archives, c(""), c("intersectionality"), span = 0.7, yearfrom = 1970, yearto = 2007)

ggsave("Figures/standardization_ceramics_overtime.png") #savefigure
```


Each keyword was then entered into it's own vector, so that I could analyze the words and the frequency of those words that are associated with each keyword. 
```{r standard_frequency_counts}

# top words per decade in articles containing 'X'
standardization_nouns <- JSTOR_dtmofnouns(multiple_archives, word = 'standardization', sparse = 1, POStag = TRUE)
#remove noisy words
custom_stopwords <- c('archaeology', 'university', 'research', 'evidence', 'journal', 'world', 'site', 'cambridge', 'archaeol', 'area', 'region', 'period', 'analysis', 'anthropology', 'springer', 'production', 'figure', 'work', 'world', 'concept', 'human', 'middle', 'altamira', 'culture', 'record', 'citation', 'discipline', 'author', 'proportion', 'literature', 'report', 'approach', 'cambridge')

standardization_freqwords <- JSTOR_freqwords(multiple_archives, standardization_nouns, custom_stopwords, n = 3)

standardization_freqwords <- JSTOR_freqwords(multiple_archives, standardization_nouns, custom_stopwords, n = 3)
```
It's clear that "standard" and "standardization" produce two different frequecy plots. From 2008-2010, pottery is the fourth word most commonly printed in association with "standard". It is also a common word during the periods of 1996-1998, the most common word associated with "standard" in the 1993-1995 literature, and the fifth most common word in articles published between 2006-2007. Standardization presents an even clearer image, in 1999-2001 article, pottery is the most common word associated with "standardization", though it falls out of favor in later years. This demonstrates that pottery and standardization/standard regularly are associated in the literature data set, suggesting a fruitful direction of comparison to the two word correlations plots presented earlier in this paper. 

```{r commodity_frequency_counts}
# top words per decade in articles containing 'X'
commodity_nouns <- JSTOR_dtmofnouns(multiple_archives, word = 'commodity', sparse = 1, POStag = TRUE)

commodity_freqwords <- JSTOR_freqwords(multiple_archives, commodity_nouns, custom_stopwords, n = 3)

```
Pottery and commodity only appear in association within articles published between 1994-1996, however, during this time period, pottery is the second most common word associated with commodity, indicating the two appear at least 400 times in conjunction. Also useful for this project is the literature published between 2006-2008, in which trade and exchange are the two most commonly associated with commodity. The question that is driving this research assumes that commodities are trasnported as a result of increased trade and exchange, therefore, the articles in which these sets of words appear together may be useful for theoretical frameworks regardless if they also contain the words pottery or ceramics. 

```{r pottery_frequency_counts}
# top words per decade in articles containing 'X'
pottery_nouns <- JSTOR_dtmofnouns(multiple_archives, word = 'pottery', sparse = 1, POStag = TRUE)

pottery_freqwords <- JSTOR_freqwords(multiple_archives, pottery_nouns, custom_stopwords, n = 3)


ceramics_nouns <- JSTOR_dtmofnouns(multiple_archives, word = 'ceramics', sparse = 1, POStag = TRUE)

ceramics_freqwords <- JSTOR_freqwords(multiple_archives, ceramics_nouns, custom_stopwords, n = 3)
```

In previous interations of this JSTORr package, it was noted that ceramics/ceramic does not always produce reliable results. Therefore, for this frequency analysis, both pottery and ceramics were utilized as key words to increase the likelihood that useful word associations could be found. It is also interesting to compare and contrast the two much the same as I did above with standard and standardization, in order to see if any differences in associations occur.  
The results of using "pottery" show that, even taking out noisey words, the data is very jumbled. The results of a frequency analysis for "ceramics" are relatively similar, possible even more jumbled. It is interesting to note that in both cases, ceramics and pottery produce too many associations to be useful, therefore, for the purpose of the paper, standardization/standard, and commodity will be utilized because they provide much more constrained data sets and are theoretically specific. 




# Introduction

The Tufte-\LaTeX\ [^tufte_latex] document classes define a style similar to the style Edward Tufte uses in his books and handouts. Tufte's style is known for its extensive use of sidenotes, tight integration of graphics with text, and well-set typography. 

# Headings

This style provides a- and b-heads (that is, `#` and `##`), demonstrated above.
An error is emitted if you try to use `###` and smaller headings.

\newthought{In his later books}[^books_be], Tufte starts each section with a bit of vertical space, a non-indented paragraph, and sets the first few words of the sentence in small caps. To accomplish this using this style, use the `\newthought` command as demonstrated at the beginning of this paragraph.

# Figures

## Margin Figures

Images and graphics play an integral role in Tufte's work. To place figures or tables in the margin you can use the `fig.margin` knitr chunk option. For example:

```{r, fig.margin = TRUE, fig.cap = "Sepal length vs. petal length, colored by species"}
library(ggplot2)
qplot(Sepal.Length, Petal.Length, data = iris, color = Species)
```

Note the use of the `fig.cap` chunk option to provide a figure caption. You can adjust the proportions of figures using the `fig.width` and `fig.height` chunk options. These are specified in inches, and will be automatically scaled down to fit within the handout margin.

## Equations

You can also include \LaTeX\ equations in the margin by explicitly invoking the `marginfigure` environment.

\begin{marginfigure}
$$\frac{d}{dx}\left( \int_{0}^{x} f(u)\,du\right)=f(x).$$
\caption{An equation}
\end{marginfigure}

Note the use of the `\caption` command to add additional text below the equation.

## Full Width Figures

You can arrange for figures to span across the entire page by using the `fig.fullwidth` chunk option. 

```{r, fig.width = 10, fig.height = 2, fig.fullwidth = TRUE, fig.cap = "Full width figure"}
qplot(wt, mpg, data=mtcars, colour=factor(cyl))
```

Note the use of the `fig.width` and `fig.height` chunk options to establish the proportions of the figure. Full width figures look much better if their height is minimized.

## Main Column Figures

Besides margin and full width figures, you can of course also include figures constrained to the main column.

```{r, fig.cap = "Another figure"}
qplot(factor(cyl), mpg, data = mtcars, geom = "boxplot")
```

# Sidenotes

One of the most prominent and distinctive features of this style is the extensive use of sidenotes. There is a wide margin to provide ample room for sidenotes and small figures. Any use of a footnote will automatically be converted to a sidenote. ^[This is a sidenote that was entered using a footnote.] 

If you'd like to place ancillary information in the margin without the sidenote mark (the superscript number), you can use the `\marginnote` command. \marginnote{This is a margin note.  Notice that there isn't a number preceding the note.}

Note also that the two footnote references (`tufte_latex` and `books_be`, both defined below) were also included in the margin on the first page of this document.

# Tables

You can use the **xtable** package to format \LaTeX\ tables that integrate well with the rest of the Tufte handout style. Note that it's important to set the `xtable.comment` and `xtable.booktabs` options as shown below to ensure the table is formatted correctly for inclusion in the document.

```{r, results='asis'}
library(xtable)
options(xtable.comment = FALSE)
options(xtable.booktabs = TRUE)
xtable(head(mtcars[,1:6]), caption = "First rows of mtcars")
```


[^tufte_latex]: https://code.google.com/p/tufte-latex/
[^books_be]: http://www.edwardtufte.com/tufte/books_be










