---
title: "Quantitative Paper"
author: "Lauryl Zenobi"
output: html_document
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, cache=TRUE, echo=FALSE)
```

#Introduction
For the purpose of this quantitative project I want to explore the relationship between several sets of words. This quantitative data will be used to improve upon the background and theoretical framework of my dissertation proposal. There is also the potential that this data will provide methodological approaches to the quantitative portion of my dissertation research. 

#Key Terms and Hypotheses
The questions that are guiding this paper consist of the following: What is the relationship between **standardization** and **ceramics** in the archaeological literature? What is the relationship between **ceramics** and **commodities**? Have these relationships changed over time? What theoretical frameworks support these types of analyses?

Due to the small sample size that is expected for these topics and study area, this project will consists of two parts. The first is a statistical analysis of the frequency of the terms identified in the preceding questions. The second part will consist of a close reading of selected articles of interest. The low numbers expected from the statistical analysis necessitate the close readings, which will be beneficial in answering higher level questions about the theoretical frameworks that were utilized in the articles of interest. 

```{r Data_ingest}
library(JSTORr)
#multiple_archives <- JSTORr ::: JSTOR_unpack_multiple_archives(mydir = "Data")
#save(multiple_archives, file="Data/multiple_archives.RData")
load("Data/multiple_archives.RData")
```

#Exploratory Data Analysis
Codes from the JSTORr package will be utilized for the statistical analysis. 
Preliminary testing of the questions above is necessary to determine which research avenue is the most likely to have a sufficiently sized data set and to produce the most useful information for the purpose of this study. I will start this process with the first potential research question (see above). The first task is to establish the frequency of key words that are used in the first research question. "Commodity", "Standardization", and "Ceramics" were analyzed for their correlation over time. 

```{r, fig.width=4, fig.height=5}
Figure_1 <- JSTOR_2wordcor(multiple_archives, "commodity", "standardization")
Figure_1$plot

Figure1word_commodity<- JSTOR_1word(multiple_archives, "commodity")
Figure1word_commodity$plot

commodity_tbl <- Figureoneword$word_by_year
commodity_tbl[which.max(commodity_tbl$word_ratio),]
commodity_tbl$word_ratio[]
#url: http://www.jstor.org/stable/279900  use last digits of doi to get to link

Figure1word_standardization<- JSTOR_1word(multiple_archives, "standardization")
Figure1word_standardization$plot
standardization_tbl <- Figure1word_standardization$word_by_year
standardization_tbl[which.max(standardization_tbl$word_ratio),]

Figure1word_standard<- JSTOR_1word(multiple_archives, "standard")
Figure1word_standard$plot
standard_tbl <- Figure1word_standard$word_by_year
standard_tbl[which.max(standard_tbl$word_ratio),] #not useful produces a coefficient comment article

Figure1word_ceramics<- JSTOR_1word(multiple_archives, "ceramics")
Figure1word_ceramics$plot
ceramics_tbl <- Figure1word_ceramics$word_by_year
ceramics_tbl[which.max(ceramics_tbl$word_ratio),]
#1593826, 20622430, tensile strength, dating shell temper, wide variety of topics.

Figure1word_ceramic<- JSTOR_1word(multiple_archives, "ceramic")
Figure1word_ceramic$plot
ceramic_tbl <- Figure1word_ceramic$word_by_year
ceramic_tbl[which.max(ceramic_tbl$word_ratio),] #Sharer and Gifford

Figure1word_pottery<- JSTOR_1word(multiple_archives, "pottery")
Figure1word_pottery$plot
pottery_tbl <- Figure1word_pottery$word_by_year
pottery_tbl[which.max(pottery_tbl$word_ratio),] #Roosevelt

#Can only get doi directly out of one word or two words function, use those articles only, supplement with other tables just for interesting things to talk about. 2wordcor and PCA doesn't give doi.
#can run the 2word cor and then the 2words function with the same words and filter them against each other to get out the articles, for example, if 2wordcor and 2words show 1989 is high year, look through those.

```
Figure 1 shows the correlation over time between the words "commodity" and "standardization".

```{r, fig.width=4, fig.height=5}
Figure_2 <- JSTOR_2wordcor(multiple_archives, "commodity", "ceramics")
Figure_2$plot #2006 is a high year
```
Figure 2 shows the correlation over time between the words "commodity" and "ceramics".

```{r, fig.width=4, fig.height=5}
Figure_3 <- JSTOR_2wordcor(multiple_archives, "standardization", "ceramics")
Figure_3$plot
```
Figure 2 shows the correlation over time between the words "standardization" and "ceramics".
  
The results of these plots indicate that while there may not be a large number of instances in which these sets of words appear together in the literature, each plot demonstrates that of the instances in which they do appear, there are a number the sets of words are strongly correlated. 

This suggests, and close readings can verify this, that the data set for this research question would be rich enough to justify a quantitative method supplemented by a close readings of selected articles. Therefore, question #1 - What is the relationships between standardization and ceramics in the archaeological literature? Has this relationship changed over time? What theoretical frameworks support these types of analyses? - will be the focus of this quantative research paper. 

To determine how each key word of interest (commodity, standardization, and ceramics/pottery) cluster within the literature, a principal component analysis of each terms was conducted. The literature data was first compiled into a vector containing only nouns, as to eliminate erroneous words. Each word was then analyzed against this vector to determine what contexts the word is utilized in within the literature. Outlier and discrete clusters of of the most interest to this research, as those will be the articles in which the terms are used uniquely. 

```{r cluster_analysis}
nouns <-JSTOR_dtmofnouns(multiple_archives)
#save(nouns, file="Data/nouns.RData")

Figure_4 <- JSTOR_clusterbywords(nouns, "commodity")
Figure_4$plot
#correlation of two vectors within range of decades
```


Each keyword was then entered into it's own vector, so that I could analyze the words and the frequency of those words that are associated with each keyword. 
```{r standard_frequency_counts}

# top words per decade in articles containing 'X'
standardization_nouns <- JSTOR_dtmofnouns(multiple_archives, word = 'standardization', sparse = 1, POStag = TRUE)

standard_nouns <- JSTOR_dtmofnouns(multiple_archives, word = 'standard', sparse = 1, POStag = TRUE)
#remove noisy words
custom_stopwords <- c('archaeology', 'university', 'research', 'evidence', 'journal', 'world', 'site', 'cambridge', 'archaeol', 'area', 'region', 'period', 'analysis', 'anthropology', 'springer', 'production', 'figure', 'work', 'world', 'concept', 'human', 'middle', 'altamira', 'culture', 'record', 'citation', 'discipline', 'author', 'proportion', 'literature', 'report', 'approach', 'cambridge')

standardization_freqwords <- JSTOR_freqwords(multiple_archives, standardization_nouns, custom_stopwords, n = 3)

standard_freqwords <- JSTOR_freqwords(multiple_archives, standard_nouns, custom_stopwords, n = 3)
```
It's clear that "standard" and "standardization" produce two different frequecy plots. From 2008-2010, pottery is the fourth word most commonly printed in association with "standard". It is also a common word during the periods of 1996-1998, the most common word associated with "standard" in the 1993-1995 literature, and the fifth most common word in articles published between 2006-2007. Standardization presents an even clearer image, in 1999-2001 article, pottery is the most common word associated with "standardization", though it falls out of favor in later years. This demonstrates that pottery and standardization/standard regularly are associated in the literature data set, suggesting a fruitful direction of comparison to the two word correlations plots presented earlier in this paper. 

```{r commodity_frequency_counts}
# top words per decade in articles containing 'X'
commodity_nouns <- JSTOR_dtmofnouns(multiple_archives, word = 'commodity', sparse = 1, POStag = TRUE)

commodity_freqwords <- JSTOR_freqwords(multiple_archives, commodity_nouns, custom_stopwords, n = 3)

```
Pottery and commodity only appear in association within articles published between 1994-1996, however, during this time period, pottery is the second most common word associated with commodity, indicating the two appear at least 400 times in conjunction. Also useful for this project is the literature published between 2006-2008, in which trade and exchange are the two most commonly associated with commodity. The question that is driving this research assumes that commodities are trasnported as a result of increased trade and exchange, therefore, the articles in which these sets of words appear together may be useful for theoretical frameworks regardless if they also contain the words pottery or ceramics. 

```{r pottery_frequency_counts}
# top words per decade in articles containing 'X'
pottery_nouns <- JSTOR_dtmofnouns(multiple_archives, word = 'pottery', sparse = 1, POStag = TRUE)

pottery_freqwords <- JSTOR_freqwords(multiple_archives, pottery_nouns, custom_stopwords, n = 3)


ceramics_nouns <- JSTOR_dtmofnouns(multiple_archives, word = 'ceramics', sparse = 1, POStag = TRUE)

ceramics_freqwords <- JSTOR_freqwords(multiple_archives, ceramics_nouns, custom_stopwords, n = 3)
```

In previous interations of this JSTORr package, it was noted that ceramics/ceramic does not always produce reliable results. Therefore, for this frequency analysis, both pottery and ceramics were utilized as key words to increase the likelihood that useful word associations could be found. It is also interesting to compare and contrast the two much the same as I did above with standard and standardization, in order to see if any differences in associations occur.  
The results of using "pottery" show that, even taking out noisey words, the data is very jumbled. The results of a frequency analysis for "ceramics" are relatively similar, possible even more jumbled. It is interesting to note that in both cases, ceramics and pottery produce too many associations to be useful, therefore, for the purpose of the paper, standardization/standard, and commodity will be utilized because they provide much more constrained data sets and are theoretically specific. 

#Selected Readings

##Theoretical Frameworks

#Discussion of Results



